---
title: "R Assignment - Admission"
author: "GWU Intro to Data Science DATS 6101"
# date: "today"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

# HW Assignment - Estimation 

This exercise uses the graduate school admissions dataset from the applicants to a higher-ed institute. The dataset is called `gradAdmit` on api.regression.fit, or as a local file `LogRegAdmit.csv`.  
The variables in the dataset are:  

* `admit`: 0 or 1 (False or True)
* `gre`: gre score
* `gpa`: grade point average at undergraduate level
* `rank`: the ranking of the applicant's undergraduate institute, 1, 2, 3, or 4


## Grad School Admissions

### Question 1  

**After importing the dataset as Adata, check the structure. Make sure admit and rank are stored as factors, instead of numeric or integers. (Hint: Use as.factor() function or factor() function).**

```{r}
Adata <- data.frame(read.csv("LogRegAdmit.csv"))
Adata
str(Adata)
Adata$admit <- as.factor(Adata$admit)
Adata$rank <- as.factor(Adata$rank)
Adata
```

Answer answer answer.



### Question 2  
**Construct t-intervals for the gre and gpa data for all applicants at 0.80 level and 0.99 level. **
```{r}
t_interval_gre_80 = t.test(x=Adata$gre, conf.level=0.80)
t_interval_gre_80
t_interval_gre_99 = t.test(x=Adata$gre, conf.level=0.99)
t_interval_gre_99
t_interval_gpa_80 = t.test(x=Adata$gpa, conf.level=0.80)
t_interval_gpa_80
t_interval_gpa_99 = t.test(x=Adata$gpa, conf.level=0.99)
t_interval_gpa_99


```
Answer answer answer. 


### Question 3  
**Repeat the same calculation for gre in Question 2 but for admitted (1) and rejected (0) separately. (You can subset them first.) Between the admitted and rejected, does the two intervals overlap at 0.80 level? And at 0.99 level?**

```{r}
admitted <- subset(Adata,admit==1)
rejected <- subset(Adata,admit==0)

t_interval_gre_80_rej = t.test(x=rejected$gre, conf.level=0.80)
t_interval_gre_80_rej 
t_interval_gre_99_rej  = t.test(x=rejected$gre, conf.level=0.99)
t_interval_gre_99_rej 
t_interval_gpa_80_rej  = t.test(x=rejected$gpa, conf.level=0.80)
t_interval_gpa_80_rej 
t_interval_gpa_99_rej  = t.test(x=rejected$gpa, conf.level=0.99)
t_interval_gpa_99_rej 

t_interval_gre_80_adm = t.test(x=admitted$gre, conf.level=0.80)
t_interval_gre_80_adm
t_interval_gre_99_adm = t.test(x=admitted$gre, conf.level=0.99)
t_interval_gre_99_adm
t_interval_gpa_80_adm = t.test(x=admitted$gpa, conf.level=0.80)
t_interval_gpa_80_adm
t_interval_gpa_99_adm = t.test(x=admitted$gpa, conf.level=0.99)
t_interval_gpa_99_adm
```

Answer answer answer. 


### Question 4   
**Make (box-) plots showing the gre distribution among applicants from different school rankings for the admitted and rejected separately. Please use ggplot for this. The x-variable should be rank, and the y-variable should be gre.**

```{r}
loadPkg("ggplot2")
ggplot(admitted, aes(x=rank, y=gre, fill=rank)) + 
  geom_boxplot() +
  labs(title="Gre distribution among applicants from different school rankings for the Admitted", x="Rankings", y = "GRE")

loadPkg("ggplot2")
ggplot(rejected, aes(x=rank, y=gre, fill=rank)) + 
  geom_boxplot() +
  labs(title="Gre distribution among applicants from different school rankings for the  rejected", x="Rankings", y = "GRE")
```

Your plot should look like the sample png image given here in the folder (for the admitted subset).
Answer answer answer. 


### Question 5   
**Repeat Question 2 and Question 3 for gpa. Do the two groups have t-intervals overlap at 0.80 level? At 0.99 level?**

```{r}
t_interval_gpa_80 = t.test(x=Adata$gpa, conf.level=0.80)
t_interval_gpa_80
t_interval_gpa_99 = t.test(x=Adata$gpa, conf.level=0.99)
t_interval_gpa_99

t_interval_gpa_80_rej  = t.test(x=rejected$gpa, conf.level=0.80)
t_interval_gpa_80_rej 
t_interval_gpa_99_rej  = t.test(x=rejected$gpa, conf.level=0.99)
t_interval_gpa_99_rej 

t_interval_gpa_80_adm = t.test(x=admitted$gpa, conf.level=0.80)
t_interval_gpa_80_adm
t_interval_gpa_99_adm = t.test(x=admitted$gpa, conf.level=0.99)
t_interval_gpa_99_adm

```

Answer answer answer. 


### Question 6   
**Repeat Question 4 for gpa. Make (box-) plots showing the gpa distribution among applicants from different school rankings for the admitted and rejected separately. Again, use ggplot.**

Answer answer answer. 
```{r}
ggplot(admitted, aes(x=rank, y=gpa, fill=rank)) + 
  geom_boxplot( ) +
  labs(title="gpa distribution among applicants from different school rankings for the admitted", x="Rankings", y = "GPA")

ggplot(rejected, aes(x=rank, y=gpa, fill=rank)) + 
  geom_boxplot( ) +
  labs(title="gpa distribution among applicants from different school rankings for the Rejected", x="Rankings", y = "GPA")
```

### Question 7  
**As EDA from the charts and the t-interval calculations you produced, do you observe any potential effects between gre/gpa/rank on admissions? Explain briefly. **

Answer answer answer.
When the rank fluctuates from rank 1 to rank 4, the boxplots of the GRE with different school rankings similarly demonstrate a minor decline in GRE and GPA results. This show that the student from Rank 4 school are performing little worst in Gre.

GRE scores and GPA both had a significant impact on admission, with the confidence intervals of the admitted group being higher than the rejected group without overlap at both 0.99 and 0.80 levels.







